`r glue::glue('inferCNV results for {params$library_id} with {params$reference_name} reference')`

### 2025-05-07 analyses/infercnv-consensus-cell-type/template-notebooks/SCPCP000015_explore-infercnv-results.Rmd
update UMAPs

umap_theme <- list(
  coord_fixed(),
  theme_classic(),
  theme(
    axis.ticks = element_blank(),
    axis.text = element_blank()
  )
)
```

@@

### CNVs in reference vs query cells

The figures below compare the number of CNVs per cell between reference and query cells.
Ideally, we'll see a bimodal distribution where reference cells are at/near zero, and there is a peak of higher values which may represent tumor cells.

@@

```{r fig.width = 12, fig.height = 5}
left_plot <- ggplot(has_cnv_df) +
  aes(y = forcats::fct_rev(cell_group_n), x = total_cnv_per_cell, fill = cell_group_n) +
  ggridges::geom_density_ridges2(alpha = 0.7, scale = 0.85) +
  labs(y = "Cell group (number of cells)") +
  theme(legend.position = "bottom")

right_plot <- ggplot(has_cnv_df) +
  aes(y = forcats::fct_rev(celltype_n), x = total_cnv_per_cell, fill = celltype_n) +
  ggridges::geom_density_ridges2(alpha = 0.7, scale = 0.85) +
  labs(y = "Cell type (number of cells)") +
  theme(legend.position = "bottom") +
  scale_fill_brewer(palette = "Dark2")

@@

### CNVs across cell types

Next, we'll look at the per-cell distributions of CNVs across cell types, using labels derived from the `cell-type-ewings` module as follows:

Ideally, we'll see that tumor and Unknown cells from will have relatively higher CNV scores compared to other cell types, in particular those cell types in the reference.

```{r fig.height=8, message=FALSE, warning=FALSE}
# drop cells from other libraries, and update data frame for plotting
query_has_cnv_df <- has_cnv_df |>
  dplyr::filter(library_id == params$library_id) |>
  dplyr::add_count(celltype) |>
  # remove really low count cell types
  dplyr::filter(n >= 3) |>
  # add column with cell type counts, and arrange in order
  dplyr::mutate(
    celltype_n = glue::glue("{celltype} (n={n})"),
    celltype_n = forcats::fct_infreq(celltype_n),
    celltype_n = forcats::fct_rev(celltype_n)
  )

ggplot(query_has_cnv_df) +
  aes(x = total_cnv_per_cell, y = celltype_n) +
  ggridges::geom_density_ridges2(alpha = 0.7, scale = 0.85)
```

@@

```{r, fig.height = 10}
celltype_umap <- ggplot(umap_df) +
  aes(x = UMAP1, y = UMAP2, color = celltype) +
  geom_point(size = 0.1, alpha = 0.3) +
  guides(color = guide_legend(override.aes = list(size = 2, alpha = 1))) +
  labs(title = "Cell types") +
  umap_theme

@@

celltype_umap / cnv_umap
```

>>>>

umap_theme <- list(
  coord_fixed(),
  theme_classic(),
  theme(
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    legend.position = "bottom"
  )
)
```

@@

The figures below compare the number of CNVs per cell between cell groupings.
Ideally, we'll see a bimodal distribution where reference cells are at/near zero, and there is a peak of higher values which may represent tumor cells.

@@

```{r fig.width = 12, fig.height = 5}
left_plot <- ggplot(has_cnv_df) +
  aes(y = forcats::fct_rev(cell_group_n), x = total_cnv_per_cell, fill = cell_group_n) +
  ggridges::geom_density_ridges2(alpha = 0.7, scale = 0.85) +
  scale_fill_brewer(palette = "Dark2") +
  labs(y = "Cell group (number of cells)") +
  theme(legend.position = "bottom")

right_plot <- ggplot(has_cnv_df) +
  aes(y = forcats::fct_rev(celltype_n), x = total_cnv_per_cell, fill = celltype_n) +
  ggridges::geom_density_ridges2(alpha = 0.7, scale = 0.85) +
  labs(y = "Cell type (number of cells)") +
  theme(legend.position = "bottom")

@@

@@

```{r, fig.height = 10}
# Filter to just cells in the library to avoid warnings
query_has_cnv_df <- has_cnv_df |>
  dplyr::filter(library_id == params$library_id)

celltype_umap <- ggplot(query_has_cnv_df) +
  aes(x = UMAP1, y = UMAP2, color = celltype_n) +
  geom_point(size = 0.1, alpha = 0.3) +
  guides(color = guide_legend(override.aes = list(size = 2, alpha = 1))) +
  labs(title = "Cell types") +
  umap_theme

@@

celltype_umap + cnv_umap
```
